---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

The questions you will have to answer are:

1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results. P.S. Should you scale? P.P.S. Don't forget to motivate your priors. P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.
1.1. Verbal IQ and ADOS
1.2. Non Verbal IQ and ADOS
1.3. Social IQ and ADOS

```{r}
# Set working directory
setwd("C:/Users/simon/GoogleDrev/Uni/Methods4/Portfolios/Portfolio2")

# Load package
library(rethinking)
library(cowplot)

#Load data
d = read.csv("Assignment3Data.csv")

# Subset data
d2 = d[d$Visit == 1,]
d3 = d2[d2$ASD == 1,]

# Scale variables
d3$VerbalIQ = scale(d3$VerbalIQ)
d3$SocialIQ = scale(d3$SocialIQ)
d3$NonVerbalIQ = scale(d3$NonVerbalIQ)

plot(d3$VerbalIQ, d3$ADOS)
plot(d3$SocialIQ, d3$ADOS)
plot(d3$NonVerbalIQ, d3$ADOS)

# Create a linear model for verbal IQ
mVerbalIQ <- map(
alist(
ADOS ~ dnorm(mu , sigma) ,
mu <- a + b*VerbalIQ,
a ~ dnorm(0, 10) ,
b ~ dnorm(0, 5) ,
sigma ~ dunif(0, 5)
) ,
data=d3)

# To get model results
precis(mVerbalIQ)

# To plot the model
ScaleVIQ.seq <- seq( from=-10 , to=15 , by=0.1 )

mu <- link( mVerbalIQ , data=data.frame(VerbalIQ=ScaleVIQ.seq) )
str(mu)
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )

sim.ados <- sim(mVerbalIQ , data=list(VerbalIQ=ScaleVIQ.seq) )
#summarizing the distribution of the simulation
ados.HPDI <- apply( sim.ados , 2 , HPDI , prob=0.89 )


# plot raw data
plot(ADOS ~ VerbalIQ , d3 , col=col.alpha(rangi2,0.5) )
# draw MAP line
lines( ScaleVIQ.seq , mu.mean )
# draw HPDI region for line
shade(mu.HPDI , ScaleVIQ.seq)
# draw PI region for simulated heights
shade(ados.HPDI , ScaleVIQ.seq)


# Create a linear model for SocialIQ
mSocialIQ <- map(
alist(
ADOS ~ dnorm(mu , sigma) ,
mu <- a + b*SocialIQ,
a ~ dnorm(0, 10) ,
b ~ dnorm(0, 5) ,
sigma ~ dunif(0, 5)
) ,
data=d3)

# To get model results
precis(mSocialIQ)

# To plot the model
ScaleSIQ.seq <- seq( from=-20 , to=30 , by=0.1 )

mu <- link( mSocialIQ , data=data.frame(SocialIQ=ScaleSIQ.seq) )
str(mu)
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )

sim.ados <- sim(mSocialIQ , data=list(SocialIQ=ScaleSIQ.seq) )
#summarizing the distribution of the simulation
ados.HPDI <- apply( sim.ados , 2 , HPDI , prob=0.89 )


# plot raw data
plot(ADOS ~ SocialIQ , d3 , col=col.alpha(rangi2,0.5) )
# draw MAP line
lines( ScaleSIQ.seq , mu.mean )
# draw HPDI region for line
shade(mu.HPDI , ScaleSIQ.seq)
# draw PI region for simulated heights
shade(ados.HPDI , ScaleSIQ.seq)

# To make a linear model of nonVerbal IQ
mNonVerbalIQ <- map(
alist(
ADOS ~ dnorm(mu , sigma) ,
mu <- a + b*NonVerbalIQ,
a ~ dnorm(0, 10) ,
b ~ dnorm(0, 5) ,
sigma ~ dunif(0, 5)
) ,
data=d3)

# To get model results
precis(mNonVerbalIQ)

# To plot the model
ScaleNVIQ.seq <- seq( from=-15 , to=20 , by=0.1 )

mu <- link(mNonVerbalIQ , data=data.frame(NonVerbalIQ=ScaleNVIQ.seq) )
str(mu)
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )

sim.ados <- sim(mNonVerbalIQ , data=list(NonVerbalIQ=ScaleNVIQ.seq) )
#summarizing the distribution of the simulation
ados.HPDI <- apply( sim.ados , 2 , HPDI , prob=0.89 )


# plot raw data
plot(ADOS ~ NonVerbalIQ , d3 , col=col.alpha(rangi2,0.5) )
# draw MAP line
lines( ScaleNVIQ.seq , mu.mean )
# draw HPDI region for line
shade(mu.HPDI , ScaleNVIQ.seq)
# draw PI region for simulated heights
shade(ados.HPDI , ScaleNVIQ.seq)

```

2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?
2.2. Build the model, assess its quality, write a few lines interpreting the results.

```{r}
dnoscale = read.csv("Assignment3Data.csv")

# Subset data
dnoscale = dnoscale[d$Visit == 1,]
dnoscale = dnoscale[d2$ASD == 1,]

dnoscale$VerbalIQ.c = dnoscale$VerbalIQ - mean(dnoscale$VerbalIQ)
dnoscale$NonVerbalIQ.c = dnoscale$NonVerbalIQ - mean(dnoscale$NonVerbalIQ)
dnoscale$SocialIQ.c = dnoscale$SocialIQ - mean(dnoscale$VerbalIQ)

cor.test(dnoscale$VerbalIQ, dnoscale$NonVerbalIQ) # 0.67
cor.test(dnoscale$NonVerbalIQ, dnoscale$SocialIQ) #0.37
cor.test(dnoscale$SocialIQ, dnoscale$VerbalIQ) #0.61

m_social=map(
  alist(
    SocialIQ ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_VerbalIQ*VerbalIQ +
        b_NonVerbalIQ*NonVerbalIQ,
    Intercept ~ dnorm(0,5),
    b_VerbalIQ ~ dnorm(0,5),
    b_NonVerbalIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d3)

precis(m_social)

# Plot residuals
mu.SIQ = coef(m_social)['Intercept'] + coef(m_social)['b_VerbalIQ']*d3$VerbalIQ + coef(m_social)['b_NonVerbalIQ']*d3$NonVerbalIQ
m.rSIQ = d3$SocialIQ - mu.SIQ

plot_SIQ = ggplot(d3, aes(y=ADOS, x= m.rSIQ)) + geom_smooth(method = 'lm') + geom_point() + xlab("SocialIQ Residuals")

m_verbal=map(
  alist(
    VerbalIQ ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_SocialIQ*SocialIQ +
        b_NonVerbalIQ*NonVerbalIQ,
    Intercept ~ dnorm(0,5),
    b_SocialIQ ~ dnorm(0,5),
    b_NonVerbalIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d3)

precis(m_verbal)

# Plot residuals
mu.VIQ = coef(m_verbal)['Intercept'] + coef(m_verbal)['b_SocialIQ']*d3$SocialIQ + coef(m_verbal)['b_NonVerbalIQ']*d3$NonVerbalIQ
m.rVIQ = d3$VerbalIQ - mu.VIQ

plot_VIQ=ggplot(d3, aes(y=ADOS, x= m.rVIQ)) + geom_smooth(method = 'lm') + geom_point() + xlab("VerbalIQ Residuals")

m_NonVerbal=map(
  alist(
    NonVerbalIQ ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_SocialIQ*SocialIQ +
        b_VerbalIQ*VerbalIQ,
    Intercept ~ dnorm(0,5),
    b_SocialIQ ~ dnorm(0,5),
    b_VerbalIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d3)

precis(m_NonVerbal)

# Plot residuals
mu.NVIQ = coef(m_NonVerbal)['Intercept'] + coef(m_NonVerbal)['b_SocialIQ']*d3$SocialIQ + coef(m_NonVerbal)['b_VerbalIQ']*d3$VerbalIQ
m.rNVIQ = d3$NonVerbalIQ - mu.NVIQ

plot_NVIQ=ggplot(d3, aes(y=ADOS, x= m.rNVIQ)) + geom_smooth(method = 'lm') + geom_point() + xlab("NonVerbalIQ Residuals")

plot_grid(plot_NVIQ, plot_SIQ, plot_VIQ)

# To get numeric estimate for residual

d3$NVIQ_res = m.rNVIQ
d3$VIQ_res = m.rVIQ
d3$SIQ_res = m.rSIQ

m_NonVerbal_res=map(
  alist(
    ADOS ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_NVIQ*NVIQ_res,
    Intercept ~ dnorm(0,5),
    b_NVIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d3)

precis(m_NonVerbal_res)

m_Verbal_res=map(
  alist(
    ADOS ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_VIQ*VIQ_res,
    Intercept ~ dnorm(0,5),
    b_VIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d3)

precis(m_Verbal_res)

m_Social_res=map(
  alist(
    ADOS ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_SIQ*SIQ_res,
    Intercept ~ dnorm(0,5),
    b_SIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d3)

precis(m_Social_res)


glimmer(ADOS ~ SocialIQ + VerbalIQ, data=d3)

m_allIQ=map(
  alist(
    ADOS ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_SocialIQ*SocialIQ +
        b_VerbalIQ*VerbalIQ +
        b_NonVerbalIQ*NonVerbalIQ,
    Intercept ~ dnorm(0,5),
    b_SocialIQ ~ dnorm(0,5),
    b_VerbalIQ ~ dnorm(0,5),
    b_NonVerbalIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d3)

precis(m_allIQ)
plot(precis(m_allIQ))

# Quality check of model

mu <- link( m_allIQ)
# summarize samples across cases
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI )

# simulate observations
# again no new data, so uses original data
ADOS.sim <- sim( m_allIQ , n=1e4 )
ADOS.HPDI <- apply( ADOS.sim , 2 , HPDI )

#plotting
plot(mu.mean ~ d3$ADOS , col=rangi2 , ylim=range(mu.HPDI) ,
xlab="Observed ADOS" , ylab="Predicted ADOS" )
abline( a=0 , b=1 , lty=2 )
for ( i in 1:nrow(d3) ) {
lines( rep(d3$ADOS[i],2) , c(mu.HPDI[1,i],mu.HPDI[2,i]) ,
col=rangi2 ) }

```


3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.

```{r}

d2=na.omit(d2)
d2$NonVerbalIQ = scale(d2$NonVerbalIQ)
d2$VerbalIQ = scale(d2$VerbalIQ)
d2$SocialIQ = scale(d2$SocialIQ)

# Create a model with all the children
m_all_child=map(
  alist(
    ADOS ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_SocialIQ*SocialIQ +
        b_VerbalIQ*VerbalIQ +
        b_NonVerbalIQ*NonVerbalIQ,
    Intercept ~ dnorm(0,5),
    b_SocialIQ ~ dnorm(0,5),
    b_VerbalIQ ~ dnorm(0,5),
    b_NonVerbalIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d2)

precis(m_all_child)

plot(precis(m_all_child))

# Quality check of model

mu <- link( m_all_child)
# summarize samples across cases
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI )

# simulate observations
# again no new data, so uses original data
ADOS.sim <- sim( m_all_child , n=1e4 )
ADOS.HPDI <- apply( ADOS.sim , 2 , HPDI )

#plotting
plot(mu.mean ~ d2$ADOS , col=rangi2 , ylim=range(mu.HPDI) ,
xlab="Observed ADOS" , ylab="Predicted ADOS" )
abline( a=0 , b=1 , lty=2 )
for ( i in 1:nrow(d2) ) {
lines( rep(d2$ADOS[i],2) , c(mu.HPDI[1,i],mu.HPDI[2,i]) ,
col=rangi2 ) }

```


4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?
4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.

5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 

```{r}
# To get numeric estimate for residual

d3$NVIQ_res = m.rNVIQ
d3$VIQ_res = m.rVIQ
d3$SIQ_res = m.rSIQ

m_NonVerbal_res=map(
  alist(
    ADOS ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_NVIQ*NVIQ_res,
    Intercept ~ dnorm(0,5),
    b_NVIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d3)

precis(m_NonVerbal_res)

m_Verbal_res=map(
  alist(
    ADOS ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_VIQ*VIQ_res,
    Intercept ~ dnorm(0,5),
    b_VIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d3)

precis(m_Verbal_res)

m_Social_res=map(
  alist(
    ADOS ~ dnorm( mu , sigma ),
    mu <- Intercept +
        b_SIQ*SIQ_res,
    Intercept ~ dnorm(0,5),
    b_SIQ ~ dnorm(0,5),
    sigma ~ dunif(0,10)
), data = d3)

precis(m_Social_res)

```